Investigating Interventions

### 2024-01-19 minos/validation/handovers.Rmd

---
params:
    file_path: '' # default arguments. specify others in command line.
    intervention: ''
title: "`r paste(params$file_path, ' ', params$intervention, ' Handover Plots')`"
output:
  #output_file: "`r paste('minos/validation/', params$file_path, '_', params$intervention, '_handovers.html') `"
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    df_print: paged
    code_folding: hide

---

@@

file_path <- params$file_path
intervention <- params$intervention

# Read raw datafiles in
raw.files <- list.files(here::here('data', 'final_US'), pattern='[0-9]{4}_US_cohort.csv', full.names = TRUE)
raw.dat <- do.call(rbind, lapply(raw.files, read.csv))

#out.path <- here::here('output', 'default_config/')
#base.dat <- read_singular_local_out(out.path, 'baseline', drop.dead = TRUE)

out.path <- here::here('output', file_path)
base.dat <- read_singular_local_out(out.path, intervention, drop.dead = TRUE)
```

@@

>>>>

---
title: "MINOS Handover Plots"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    df_print: paged
    code_folding: hide
---

@@

```{r}
# Read raw datafiles in
raw.files <- list.files(here::here('data', 'final_US'), pattern='[0-9]{4}_US_cohort.csv', full.names = TRUE)
raw.dat <- do.call(rbind, lapply(raw.files, read.csv))

out.path <- here::here('output', 'default_config/')
base.dat <- read_singular_local_out(out.path, 'baseline', drop.dead = TRUE)
```

@@

## Hourly Wage

```{r}
handover_boxplots(raw.dat, base.dat, 'hourly_wage')
handover_lineplots(raw.dat, base.dat, 'hourly_wage')
```

```{r}
raw.s <- raw.dat %>%
  select(pidp, age, time, hourly_wage)
base.s <- base.dat %>%
  select(pidp, age, time, hourly_wage)

spaghetti_plot(spag, 'hourly_wage',
               save = shall.we.save,
               save.path = save.path)

# now a version limited to 0-100
spag <- spag %>%
  filter(hourly_wage <= 300)
spaghetti_plot(spag, 'hourly_wage',
               save = shall.we.save,
               save.path = save.path)

```{r}
# some basic statistics about hourly_wage variable
print("Some basic statistics on hourly_wage. These are all for 2020.")
raw.2020 <- raw.dat %>% filter(time == 2020)
print('All living people in 2020')
print(paste0("Number of people alive in 2020: ", nrow(raw.2020)))
# number individuals with an hourly_wage > 0
print('People not in work')
print(paste0("Number of individuals not in paid work: ", sum(!raw.2020$S7_labour_state %in% c('FT Employed', 'PT Employed'))))
print(paste0("Number of individuals with hourly_wage == 0: ", sum(raw.2020$hourly_wage == 0)))

# Number in work
print('People in work')
print(paste0("Number of individuals in paid work: ", sum(raw.2020$S7_labour_state %in% c('FT Employed', 'PT Employed'))))
print(paste0("Number of individuals with hourly_wage > 0: ", sum(raw.2020$hourly_wage > 0)))

# Extreme hourly_wage values
print('Extreme hourly wage values: HIGH')
print(paste0("Number of individuals with hourly_wage > 30: ", sum(raw.2020$hourly_wage > 30)))
print(paste0("Number of individuals with hourly_wage > 50: ", sum(raw.2020$hourly_wage > 50)))
print(paste0("Number of individuals with hourly_wage > 100: ", sum(raw.2020$hourly_wage > 100)))
print(paste0("Number of individuals with hourly_wage > 500: ", sum(raw.2020$hourly_wage > 500)))
print(paste0("Number of individuals with hourly_wage > 1000: ", sum(raw.2020$hourly_wage > 1000)))
print(paste0("Number of individuals with hourly_wage > 2000: ", sum(raw.2020$hourly_wage > 2000)))

raw.2020.work <- raw.2020 %>% filter(hourly_wage > 0)
print('Extreme hourly wage values: LOW (non-zero)')
print(paste0("Number of individuals with hourly_wage < 20: ", sum(raw.2020.work$hourly_wage < 20)))
print(paste0("Number of individuals with hourly_wage < 10: ", sum(raw.2020.work$hourly_wage < 10)))
print(paste0("Number of individuals with hourly_wage < 5: ", sum(raw.2020.work$hourly_wage < 5)))
print(paste0("Number of individuals with hourly_wage < 1: ", sum(raw.2020.work$hourly_wage < 1)))
```

## S7 Labour State

```{r}
prepped.list <- validation_prep_ordinal(raw.dat, base.dat, 'S7_labour_state')

lab.state <- prepped.list$var
lab.state.norm <- prepped.list$norm

# Now plot
ggplot(data = lab.state, mapping = aes(x = time, y = n, group = S7_labour_state, colour = S7_labour_state)) +
  geom_line() +
  geom_vline(xintercept=start.year, linetype='dotted') +
  labs(title = 'S7 Labour State', subtitle = 'Counts by Level') +
  xlab('Year') +
  ylab('Count')

ggplot(data = lab.state.norm, mapping = aes(x = time, y = prct, fill=S7_labour_state)) +
  geom_bar(stat = 'identity') +
  geom_vline(xintercept=start.year, linetype='dotted') +
  labs(title = 'S7 Labour State') +
  xlab('Year') +
  ylab('Proportion')

if (shall.we.save) {
  ggsave(filename = 'S7_labour_state.png',
         plot = last_plot(),
         path = save.path)
}

rm(prepped.list, lab.state, lab.state.norm)
```

```{r}
pivoted <- combine_and_pivot_long(df1 = base.dat, 
                                       df1.name = 'simulated', 
                                       df2 = raw.dat, 
                                       df2.name = 'raw', 
                                       var = 'S7_labour_state')

cv_ordinal_plots(pivoted.df = pivoted, 
                 var = 'S7_labour_state',
                 save = shall.we.save,
                 save.path = save.path)
```

## Child Ages

```{r}
raw.child.ages <- raw.dat %>%
  dplyr::select(pidp, hidp, time, "child_ages", "nkids") %>%
  filter(time > 2013) %>%
  group_by(time, hidp) %>%
  summarise(child_ages = find_mode(child_ages))

r.s<- strsplit(raw.child.ages$child_ages, "_")

raw.child.ages.expanded <- data.frame(time = rep(raw.child.ages$time, sapply(r.s, length)), expanded_child_ages = unlist(r.s))
raw.child.ages.expanded <- raw.child.ages.expanded %>%
  group_by(time, expanded_child_ages) %>%
  summarise(age_counts = n()) %>%
  mutate(source = 'final_US') %>%
  filter(expanded_child_ages <= 16)

base.child.ages <- base.dat %>%
  dplyr::select(pidp, hidp, time, "child_ages", "nkids") %>%
  filter(time > 2009) %>%
  group_by(time, hidp) %>%
  summarise(child_ages = find_mode(child_ages))

b.s<- strsplit(base.child.ages$child_ages, "_")

base.child.ages.expanded <- data.frame(time = rep(base.child.ages$time, sapply(b.s, length)), expanded_child_ages = unlist(b.s))
base.child.ages.expanded <- base.child.ages.expanded %>%
  group_by(time, expanded_child_ages) %>%
  summarise(age_counts = n()) %>%
  mutate(source = 'baseline')

combined <- rbind(raw.child.ages.expanded, base.child.ages.expanded)

ggplot(data = combined, mapping = aes(x = time, y = age_counts, group = expanded_child_ages, colour =  factor(as.numeric(expanded_child_ages)))) +
  geom_line() +
  geom_vline(xintercept = start.year) +
  labs(title = "Child Age Counts Over Time", subtitle = 'Counts by Level') +
  scale_colour_viridis_d() +
  xlab('Year') +
  ylab('Count')

ggplot(data = combined, mapping = aes(x = factor(as.numeric(expanded_child_ages)), y = age_counts, group = factor(time), colour = factor(time))) +
  geom_line() +
  geom_vline(xintercept = start.year) +
  labs(title = "Child Age Counts Over Time", subtitle = 'Counts by Level') +
  scale_colour_viridis_d() +
  xlab('Year') +
  ylab('Count')

base.child.ages.nkids <- base.dat %>%
  group_by(time, hidp) %>%
  summarise(nkids_counts = mean(nkids)) %>%
  group_by(time) %>%
  summarise(nkids_by_year = sum(nkids_counts)) %>%
  mutate(source = 'baseline')

raw.child.ages.nkids <- raw.dat %>%
  group_by(time, hidp) %>%
  summarise(nkids_counts = mean(nkids)) %>%
  group_by(time) %>%
  summarise(nkids_by_year = sum(nkids_counts)) %>%
  mutate(source = 'final_US')

combined.nkids <- rbind(base.child.ages.nkids, raw.child.ages.nkids)

ggplot(data = combined.nkids, mapping = aes(x = time, y = nkids_by_year, group = source, colour = source)) +
  geom_line() +
  labs(title = "nkids Count Over Time", subtitle = 'Counts by Level') +
  xlab('Year') +
  ylab('Count')

### 2024-01-19 minos/validation/interventions.Rmd

>>>>

## Poverty Line Child Uplift

```{r}
income.base <- base %>% 
  select(pidp, time, hh_income)
income.poverty <- poverty %>% 
  select(pidp, time, hh_income, income_boosted)
# first rename and cbind
income.b <- rename(income.base, base = hh_income)
income.e <- rename(income.poverty, poverty = hh_income)
income.merged <- merge(income.b, income.e, by=c('pidp', 'time'))
income.merged <- pivot_longer(data = income.merged,
                              cols = base:poverty,
                              names_to = 'scenario',
                              values_to = 'hh_income')
income.merged.byyr <- income.merged %>%
  group_by(time, scenario) %>%
  summarise(hh_income = mean(hh_income))
ggplot(data = income.merged.byyr, mapping = aes(x = time, y = hh_income, color=scenario)) +
  geom_line()
```
## SF_12
```{r}
sf12.base <- base %>% 
  select(pidp, time, SF_12)
sf12.poverty <- poverty %>% 
  select(pidp, time, SF_12, income_boosted)
# first rename and cbind
sf12.b <- rename(sf12.base, base = SF_12)
sf12.e <- rename(sf12.poverty, poverty = SF_12)
sf12.merged <- merge(sf12.b, sf12.e, by=c('pidp', 'time'))
sf12.merged <- pivot_longer(data = sf12.merged,
                              cols = base:poverty,
                              names_to = 'scenario',
                              values_to = 'SF_12')
sf12.merged.byyr <- sf12.merged %>%
  group_by(time, scenario) %>%
  summarise(SF_12 = mean(SF_12))
ggplot(data = sf12.merged.byyr, mapping = aes(x = time, y = SF_12, color=scenario)) +
  geom_line()
```

## Boost Amount
```{r}
income.poverty <- poverty %>% 
  select(pidp, time, hh_income, income_boosted, boost_amount)
income.poverty.byyr <- income.poverty %>%
  group_by(time) %>%
  summarise(boost_amount = mean(boost_amount))
ggplot(data = income.poverty.byyr, mapping = aes(x = time, y = boost_amount)) +
  geom_line()
```
## Who Boosted?

```{r}
boosted.byyr <- income.poverty %>% 
  group_by(time) %>%
  count(income_boosted)
ggplot(data = filter(boosted.byyr, income_boosted == 'True'), mapping = aes(x = time, y = n)) +
  geom_line()
```
